---
title: "Tutorial: How to use fhpredict"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Install the package

Install the package fhpredict from GitHub, using the package remotes. Install
from the "dev" branch to get the latest version:

```{r eval = FALSE}
# install.packages(remotes)

remotes::install_github("kwb-r/fhpredict@dev", build_vignettes = TRUE)
```

## Users

In the data model of the app everything is assigned to a user identified by
a unique identifier, the user id. To get an overview on available users and 
their associated ids, run:

```{r}
# Get overview on available users in a data frame
users <- fhpredict::api_get_users()

# Print the data frame
users
```

## Models

### Overview on available models

Use the function `api_get_model()` to get an overview on the models that are
stored in the postgres database. You need to pass the user id (here: 3) and
the id of the bathing spot (here: 18) to the function.

```{r}
# Read all models that are stored for one bathing spot of one user
model_info <- fhpredict::api_get_model(user_id = 3, spot_id = 18)
```

The function returns a data frame with one row per available model:

```{r}
# Show meta information on the models
model_info
```

Use this overview on available models to lookup the id of the model that you 
actually want to fetch from the database. 

### Fetching a model

Use the model id to fetch a specific model from the database:

```{r}
# Load model with id 27 from the database
model <- fhpredict::api_get_model(user_id = 3, spot_id = 18, model_id = 27)
```

```{r}
# Show the (rstan) model
print(model)
```

### Saving a model

For testing purposes we store a simple, small object instead of a STAN model.
We use the `cars` dataset that is shipped with "base R". Use the function 
`api_add_model()` to add the "model" to the database:

```{r}
# Look at the head of the cars dataset
head(cars)

# Add the cars dataset to the database
model_id <- fhpredict:::api_add_model(
  user_id = 3, spot_id = 18, model = cars, comment = "Cars in R"
)
```

The function returns the id of the model that was given by the database. We
stored the id in the variable `model_id`. To check if the model arrived in the
database we read it back, again using `api_get_model()`

```{r}
# Read the "model" back
my_cars <- fhpredict::api_get_model(3, 18, model_id)
```

We convince ourselves that what we get is identical to what we stored:

```{r}
# Nothing should have changed!
identical(cars, my_cars)
```

### Deleting a model

Use the function `api_delete_model()` to remove a model from the database. 
Let's delete the model that we just added. Its id is given in the variable
`model_id`.

```{r}
# Delete the model with the id given in "model_id"
fhpredict::api_delete_model(3, 18, model_id)
```

## Purification Plants

Use the following script to check if there are any purification plants defined.
As there are many bathing spots, the script takes quite a long time and is not
run here so that you do not see any outputs.

```{r eval = FALSE}
# Get meta information on all bathing spots
bathing_spots <- fhpredict::api_get_bathingspot(user_id = 3, limit = 10000)

# Number of bathing spots
nrow(bathing_spots)

# Access the "purificationPlants" endpoint for each bathing spot
purification_plants <- lapply(bathing_spots$id, function(spot_id) {
  
  message("spot id = ", spot_id)
  
  result <- fhpredict::postgres_get(path = sprintf(
    "%s/purificationPlants", 
    fhpredict:::path_bathingspot(user_id = 3, spot_id = spot_id)
  ))
  
  fhpredict:::stop_on_request_failure(result)
  
  result$data
})

# How many bathing spots have purification plants?
sum(lengths(purification_plants) > 0)
```

