---
title: "Tutorial: How to use fhpredict"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Install the package

Install the package fhpredict from GitHub, using the package remotes. Install
from the "dev" branch to get the latest version:

```{r eval = FALSE}
# install.packages(remotes)

remotes::install_github("kwb-r/fhpredict@dev")
```

## Models

### Overview on available models

Use the function `api_get_model()` to get an overview on the models that are
stored in the postgres database. You need to pass the user id (here: 3) and
the id of the bathing spot (here: 18) to the function.

```{r}
# Read all models that are stored for one bathing spot of one user
models <- fhpredict::api_get_model(user_id = 3, spot_id = 18)
```

The function returns a list with each list element representing one model. 

The metadata about the models are returned in the attribute "model_info" of the
returned list:

```{r}
# Get and show meta information on the models
(model_info <- kwb.utils::getAttribute(models, "model_info"))
```

Use the list of available models to lookup the id of the model that you want
to fetch from the database. 

### Fetching a model

Use the model id to fetch a specific model from the database:

```{r}
# Load model with id 36 from the database
model <- fhpredict::api_get_model(user_id = 3, spot_id = 18, model_id = 36)

# Show the model
model
```

### Saving a model

For testing purposes we store a simple, small object instead of a STAN model.
We use the `cars` dataset that is shipped with "base R". Use the function 
`api_add_model()` to add the "model" to the database:

```{r}
# Look at the head of the cars dataset
head(cars)

# Add the cars dataset to the database
model_id <- fhpredict:::api_add_model(
  user_id = 3, spot_id = 18, model = cars, comment = "Cars in R"
)
```

The function returns the id of the model that was given by the database. We
stored the id in the variable `model_id`. To check if the model arrived in the
database we read it back, again using `api_get_model()`

```{r}
# Read the "model" back
my_cars <- fhpredict::api_get_model(3, 18, model_id)
```

We convince ourselves that what we get is identical to what we stored:

```{r}
# Nothing should have changed!
identical(cars, my_cars)
```

### Deleting a model

Use the function `api_delete_model()` to remove a model from the database. 
Let's delete the model that we just added. Its id is given in the variable
`model_id`.

```{r eval}
# Delete the model with the id given in "model_id"
fhpredict::api_delete_model(3, 18, model_id)
```

